OpReg-Boost: Learning to Accelerate Online Algorithms \\ with Operator Regression

### 2021-01-14 paper/main.tex
small things
Third pass

%-------------------- COMMANDS
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\minimize}{minimize}
\DeclareMathOperator{\prox}{prox}
\DeclareMathOperator{\fix}{fix}

@@

% new math commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\f}{\mathbold{f}}
\newcommand{\bdelta}{\mathbold{\delta}}
\newcommand{\transp}{\top}
\newcommand{\bv}{\mathbold{b}}
\newcommand{\e}{\mathbold{e}}
\newcommand{\p}{\mathbold{p}}
\newcommand{\q}{\mathbold{q}}
\newcommand{\tv}{\mathbold{t}}
\newcommand{\x}{\mathbold{x}}
\newcommand{\y}{\mathbold{y}}
\newcommand{\vv}{\mathbold{v}}
\newcommand{\z}{\mathbold{z}}
\newcommand{\xx}{\pmb{\xi}}

\newcommand{\Am}{\mathbold{A}}
\renewcommand{\Im}{\mathbold{I}}
\newcommand{\Pm}{\mathbold{P}}

\newcommand{\I}{\mathcal{I}}
\newcommand{\T}{\mathcal{T}}

@@

\newcommand{\nicola}[1]{{\color{blue}#1}}

@@

\begin{abstract}
We present OpReg-Boost, a novel acceleration scheme to boost the convergence properties and lessen the asymptotical error of online algorithms for time-varying (weakly) convex optimization problems. OpReg-Boost is built to learn the closest algorithm to a given online algorithm that has all the algorithmic properties one needs for fast convergence, and it is based on the concept of operator regression. We show how to build OpReg-Boost by using a Peaceman-Rachford solver, and further boost [... interpolation-MAP..]. Simulation results showcase the [better/increase/..] properties of OpReg-Boost w.r.t. the more classical [...], and its close relative convex-regression-boost which is also novel but significantly less performing.  
\end{abstract}

@@

[Add others' papers, esp. ML community]

@@

Solving time-varying optimization problems with online algorithms amounts to find and track continuously changing optimizers. If we let the optimizers trajectory be defined as $\x^*(t)$, a significant results for online algorithms in this context is that if the \emph{path-length} $\sum_{k=1}^T \|\x^*(t_{k+1}) - \x^*(t_{k})\|$, for subsequent time instances $t_{k+1}, t_k$, is not sub-linear in $T$, then online algorithms will have an asymptotical error $O(1)$~\cite{Besbes2013, Jadbabaie2015, SPM, NaLi2020} (or $O(T)$ in terms of pertinent notions of regret)\footnote{
%%% Footnote Jadbabaie:
In~\cite{Jadbabaie2015} and subsequent works, [Finish]
%%%
}. While errors can be made small, for instance with the use of prediction-correction algorithms~\cite{Dontchev2013, simonetto_time-varying_2020}, the existence of an asymptotical error remains a distinctive feature of online algorithms for time-varying optimization.

@@

This gives rise to convex-regression-based boosting.

@@

In this paper, we offer the following contributions,
\begin{enumerate}
\item We present two novel regression methods to \emph{learn-project-and-solve} time-varying optimization problems that are not necessarily convex. The methods are based on convex regression and operator regression, and -- especially the latter -- is promising in boosting convergence without introducing a larger asymptotical error. 
%%
\item We present efficient ways to solve the operation regression problem via a pertinent reformulation of the Peaceman-Rachford splitting method. [.. more ]
%%
\item We [Interpolation]
%%
\item We showcase the [simulation results]
\end{enumerate}

@@

@@

The acceleration schemes that we compare with are [Tell and cite] e.g., AA: \cite{pmlr-v119-mai20a}

@@

The prox operator [..]

@@

We are interested in solving the following time-varying optimization problem,
\begin{equation}\label{eq:continuous-time-problem}
	\x^*(t) \in \argmin_{\x} f(\x; t) + g(\x; t)
\end{equation}
where $f : \R^n \times \R_+ \to \R$ is closed, proper, and $\mu$-weakly convex, whereas $g : \R^n \times \R_+ \to \R \cup \{ +\infty \}$ is closed, convex and proper function, optionally with $g \equiv 0$. Upon sampling the problem at discrete time instances $t_k$, we then obtain the sequence of time-invariant problems,
\begin{equation}\label{eq:base-problem}
	\x^*(t_k) \in \argmin_{\x} f(\x; t_k) + g(\x; t_k), \qquad k \in \N. 
\end{equation}
Our overarching goal is to set up an online algorithm $\mathcal{A}$ that generate a sequence of approximate optimizers $\{\x_k\}_{k\in \N}$, such that the asymptotical tracking error 
\begin{equation}
\limsup_{k \to \infty} \|\x_k - \x^*(t_k)\|,
\end{equation}
is as small as possible. Our main blanket assumption is that optimizers\footnote{..} at subsequent time instances cannot be arbitrarily far apart, that is, 
\begin{equation}
\|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta < +\infty, \qquad k \in \N,
\end{equation}
which in turn guarantees that the path-length grows linearly in time, 
\begin{equation}
\sum_{k\in \N} \|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta k.
\end{equation}

The first important and known result here is that if function $f$ is in $\mathcal{S}_{m,L}(\R^n)$ uniformly in $k$, then an online algorithm $A$ (here a forward-backward algorithm) can obtain linear convergence to the asymptotical error bound $\Delta/(1-\gamma)$, with optimal $\gamma = \frac{L-m}{L+m}$; much weaker results hold for the general convex case~\cite{Simonetto20XX}, for example [..]

The second important result is that even a simple Tikhonov regularization $+ \frac{w}{2} \|x\|^2_2$, transforming a smooth but not strongly convex case into a strongly convex one allows for a boost in convergence and [..]

%% say something more, change?

@@

@@

We briefly introduce here the concept of convex regression, while we leave the main technical details to~\cite{Mazumder2019,simonetto_smooth_2021}. Suppose one has collected $\ell$ noisy measurements of a convex function $\varphi(\x): \R^n \to \R$ (say $y_i$) at points $\x_i \in \R^n$, $i \in I_{\ell}$. Then convex regression is a least-squares approach to estimate the generating function based on the measurements. Formally, letting $\varphi \in \mathcal{S}_{m,L}(\R^n)$, then one would like to solve the infinite-dimensional problem, 
\begin{equation}\label{eq.inf}
\hat{\varphi}_{\ell} \in \argmin_{\psi \in \mathcal{S}_{m, L}(\R^n)}\Big\{ \sum_{i\in I_{\ell}} (y_i - \psi(\x_i))^2 \Big\} \,.
\end{equation}

The problem can be then equivalently decomposed into an estimation on the data points, to find the true function values and gradients at the data point ($\f = [\varphi_i]_{i \in I_{\ell}}$, $\bdelta = [\nabla \varphi_i]_{i \in I_{\ell}}$) which turns out to be a convex quadratically constrained quadratic program, 
\begin{subequations}\label{socp}
\begin{eqnarray}
(\f^*, \bdelta^*) &&= \argmin_{\f\in\R^\ell\!,~ \bdelta \in\R^{n\ell}} \sum_{i\in I_{\ell}} (y_i - \varphi_i)^2 \\
\mathrm{s.t.:} &&
\varphi_i - \varphi_j - \bdelta_j^\transp (\x_i - \x_j) \geq \\ && \hskip-5mm \frac{1}{2(1 - m/L)}\left(\frac{1}{L}\|\bdelta_i - \bdelta_j \|^2_2 + m \|\x_i - \x_j\|_2^2 \right. \nonumber \\ && \hskip-5mm \left. - 2 \frac{m}{L} (\bdelta_j-\bdelta_i)^\transp (\x_j - \x_i) \right), \quad \forall i, j \in I_{\ell}. \nonumber
\end{eqnarray} 
\end{subequations}

@@

Then, to solve Q1, one could consider a number of evaluations of the function $f(\x; t_k)$ as noisy measurements of an underlying convex function $\varphi_k \in \mathcal{S}_{m,L}(\R^n)$ and using the least-squares approach above to \emph{project} function $f(\cdot; t_k)$ onto the space of ``good'' functions [Complete.. put the details in the appendix]

@@

A slightly different approach to convex regression is to look at the algorithm directly, instead of the function. First of all, we remind the reader that any algorithm can be seen as an operator that maps the current approximate optimizer $\x_k$ into a new approximate optimizer $\x_{k+1}$. For example, a gradient algorithm of the form,
\begin{equation}
\x_{k+1} = \x_k - \alpha \nabla_{\x} f(\x_k; t_k) \equiv \underbrace{(I - \alpha \nabla_{x}f_k)}_{=:\T_k} (\x_k)
\end{equation}
where $\T_k: \R^n \to \R^n$ is the gradient algorithm operator. Interpreting algorithms as operators (possibly averaged, monotone, etc.) has been extremely fruitful for characterizing their convergence properties~\cite{Rockafellar1976, Eckstein1989, bauschke_convex_2017,Ryu2015,tom-also}.

@@

First of all, using Fact~2.2 in \cite{ryu_operator_2020}, we know that an operator $\T$ is $\zeta$-contractive interpolable (and therefore extensible to the whole space) if and only if it satisfies
\begin{equation}
	\norm{\T \x_i - \T \x_j}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2, \quad \forall i,j \in I_{\ell}, \ i \neq j.
\end{equation}
Therefore, we can define the following convex quadratically constrained quadratic program as our regression problem:
\begin{equation}\label{eq:operator-regression}
\begin{split}
	\hat{\tv} &= \argmin_{ \R^{n\ell} \ni \tv = [\tv_i]_{i \in I_{\ell}}} \frac{1}{2} \sum_{i \in I_{\ell}} \norm{\tv_i - \T_k \x_i}^2 \\
	&\text{s.t.} \ \norm{\tv_i - \tv_j}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2 \ \forall i,j \in I_{\ell}, i \neq j,
\end{split}
\end{equation}
where again the cost function is a least square terms on the ``observations''' and the constraints enforce contractiveness. In particular, the optimal values $\hat{\tv}$ on the data points represent the evaluations of a $\zeta$-contracting operator when applied to those points.

To extend [interpolation here]

%% Then Interpolation here..

@@

@@

[Add here]

@@

Therefore, we are interested in learning a contracting $\hat{\T}_k$ by evaluations of $\T_k$. The OpReg-Boost algorithm can be described as follows: at each time $t_k$ do:
\begin{enumerate}
	\item Sample a new problem, that is, observe $f(\x; t_k)$ and $g(\x; t_k)$;

	\item Learn the closest contracting operator to $\T_{k}$, say $\hat{\T}_k$ and

	\item Apply $\x_{k+1} = \prox_{g_k} (\hat{\T}_k \x_k)$.
\end{enumerate}
The learning step is performed using the (novel) constrained operator regression described in~\eqref{.}, specifically:

\begin{itemize}
	\item Choose $\ell$ points around $\x_k$ including $\x_k$ itself (e.g., adding a zero-mean Gaussian noise term)

	\item Evaluate the operator on the data points: $\tv_i = \T_{k} \x_i$, $i\in I_{\ell}$, for example $\tv_i = \x_i - \alpha \nabla_{\x} f(\x_i; t_k)$;

	\item Solve~\eqref{eq:operator-regression} with the PRS-based solver, and output $\hat{\tv}_k = \hat{\T}_k \x_k$ to be used in Step 3. of OpReg-Boost.
\end{itemize}

[Some closing words, no need for interpolation? ]

@@

[Add here]

@@

>>>>

%-------------------- COMMANDS
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\minimize}{minimize}
\DeclareMathOperator{\prox}{prox}
\DeclareMathOperator{\fix}{fix}
\DeclareMathOperator{\proj}{proj}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{remark}[theorem]{Remark}

@@

% new math commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\f}{\mathbold{f}}
\newcommand{\bdelta}{\mathbold{\delta}}
\newcommand{\transp}{\top}
\newcommand{\bv}{\mathbold{b}}
\newcommand{\e}{\mathbold{e}}
\newcommand{\p}{\mathbold{p}}
\newcommand{\q}{\mathbold{q}}
\newcommand{\tv}{\mathbold{t}}
\newcommand{\x}{\mathbold{x}}
\newcommand{\y}{\mathbold{y}}
\newcommand{\vv}{\mathbold{v}}
\newcommand{\z}{\mathbold{z}}
\newcommand{\xx}{\pmb{\xi}}
\newcommand{\ball}[2]{\mathbb{B}_{#2}(#1)}
\newcommand{\Am}{\mathbold{A}}
\renewcommand{\Im}{\mathbold{I}}
\newcommand{\Pm}{\mathbold{P}}
\newcommand{\cv}{\mathbold{c}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\T}{\mathcal{T}}

@@

\newcommand{\nicola}[1]{{\color{blue}#1}}
\newcommand{\andrea}[1]{{\color{red}[#1]}}

@@

\begin{abstract}
We present OpReg-Boost, a novel acceleration scheme to boost the convergence properties and lessen the asymptotical error of online algorithms for time-varying (weakly) convex optimization problems. OpReg-Boost is built to learn the closest algorithm to a given online algorithm that has all the algorithmic properties one needs for fast convergence, and it is based on the concept of operator regression. We show how to compute OpReg-Boost by using a Peaceman-Rachford solver, and further trade-off computations and accuracy with an interpolation-based technique based on alternating projection. Simulation results showcase the \andrea{better/increase/..} properties of OpReg-Boost w.r.t. the more classical \andrea{...}, and its close relative convex-regression-boost (CvxReg-Boost) which is also novel but significantly less performing.  
\end{abstract}

@@

\andrea{Add others' papers, esp. ML community}

\andrea{Notation to iron}

@@

Solving time-varying optimization problems with online algorithms amounts to find and track continuously changing optimizers $\x^*(t)$ and optimal values $F^*(t)$. A significant results for discrete-time online algorithms, that sample and solve sequences of time-varying problems at discrete time-instances $t_k$, $k \in N$, is that if the variations of the problem in time are not vanishing, then online algorithms will have an asymptotical error in terms of optimality, depending on those variations~\cite{Besbes2013, Jadbabaie2015, SPM, NaLi2020}\footnote{
%%% Footnote Jadbabaie:
In~\cite{Jadbabaie2015} and subsequent works, these asymptotical error is formulated in terms of a pertinent notion of regret, and the variations in terms of path-length and functional variability.
%%%
}. While errors can be made small, for instance with the use of prediction-correction algorithms~\cite{Dontchev2013, simonetto_time-varying_2020}, the existence of an asymptotical error remains a distinctive feature of online algorithms for time-varying optimization.

@@

This gives rise to convex-regression-based boosting, or CvxReg-Boost.

@@

In this paper, we offer the following contributions,
\begin{enumerate}
\item We present two novel regression methods to \emph{learn-project-and-solve} time-varying optimization problems that are not necessarily convex. The methods are based on convex regression and operator regression, and -- especially the latter -- is promising in boosting convergence without introducing a larger asymptotical error. 
%%
\item We present efficient ways to solve the regression problems via a pertinent reformulation of the Peaceman-Rachford splitting (PRS) method. Our PRS method is trivially parallel and allows for a reduction of the per iteration complexity from $O()$ to $O()$, where \andrea{put in the numbers}.  
%%
\item We show how to make use of an 1945 interpolation technique to trade-off accuracy and speed in operation regression, by interpolating the learned operator outside the data points via alternating projections. 
%%
\item We showcase the \andrea{simulation results}
\end{enumerate}

@@

\andrea{add ML work}

@@

The acceleration schemes that we compare with are \andrea{Tell and cite, e.g., AA: \cite{pmlr-v119-mai20a}}

@@

We indicate with $\prox_{\alpha g}(\cdot)$ the proximal operator of function $g$, and with $\proj_{C}(\cdot)$, the projection operator onto the set $C$.

@@

We are interested in solving the following time-varying optimization problem,
\begin{equation}\label{eq:continuous-time-problem}
	\x^*(t) \in \argmin_{\x} f(\x; t) + g(\x; t) =: F(\x; t)
\end{equation}
where $f : \R^n \times \R_+ \to \R$ is closed, proper, and $\mu$-weakly convex, whereas $g : \R^n \times \R_+ \to \R \cup \{ +\infty \}$ is closed, convex and proper function, optionally with $g \equiv 0$. Upon sampling the problem at discrete time instances $t_k$, we then obtain the sequence of time-invariant problems,
\begin{equation}\label{eq:base-problem}
	\x^*(t_k) \in \argmin_{\x} f(\x; t_k) + g(\x; t_k), \qquad k \in \N. 
\end{equation}
Our overarching goal is to set up an online algorithm of the forward-backward type,
\begin{equation}
\x_{k+1} = \prox_{\alpha g(\cdot, t_k)} (\x_k - \alpha \nabla_{\x} f(\x_k; t_k)), \quad k \in \N,
\end{equation}
that generate a sequence of approximate optimizers $\{\x_k\}_{k\in \N}$, such that the asymptotical tracking error \andrea{
\begin{equation}
\limsup_{k \to \infty} \|\x_k - \x^*(t_k)\|,
\end{equation}
is as small as possible. Our main blanket assumption is that optimizers\footnote{..} at subsequent time instances cannot be arbitrarily far apart, that is, 
\begin{equation}
\|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta < +\infty, \qquad k \in \N,
\end{equation}
which in turn guarantees that the path-length grows linearly in time, 
\begin{equation}
\sum_{k\in \N} \|\x^*(t_{k+1}) - \x^*(t_k)\| \leq \Delta k.
\end{equation}

Here should we formulate it in terms of functional $|f - f^*|$ instead? Let's discuss.

Somewhere we have to say that forward-backward may not converge in general, while here we ensure convergence since we are convex. This is an important point

}

\andrea{Explain a bit better what we mean by online-- I think we could be clearer?? Or maybe in the introduction.}

\andrea{Is $f$ differentiable, does it have to be?}

The first important and known result here is that if function $f$ is in $\mathcal{S}_{m,L}(\R^n)$ uniformly in $k$, then an online forward-backward algorithm can obtain linear convergence to the asymptotical error bound $\limsup_{k\to\infty}\|\x_k - \x^*(t_k)\| \leq \gamma\Delta/(1-\gamma)$, with optimal $\gamma = \frac{L-m}{L+m}$; much weaker results hold for the general convex case~\cite{Simonetto20XX}.

The second important result is that even a simple Tikhonov regularization $+ \frac{w}{2} \|x\|^2_2$, transforming the above $f \in \mathcal{S}_{m,L}(\R^n)$ to $f' \in \mathcal{S}_{m',L'}(\R^n)$ with a better $\gamma'<\gamma$, could allow for a boost in convergence and reducing the asymptotical error. In fact, if the optimizers of the true function $F = f+g$ and regularized function $F' = f'+g$ are never farther apart than $e_{\x}$, then the asymptotical error for an algorithm on $F'$ is $\gamma'\Delta'/(1-\gamma') + e_{\x}$, which can be less than the original $\gamma\Delta/(1-\gamma)$.

\andrea{say something more, change above -- these are worst case bounds..?}

@@

In the following, we will explore both Q1 and Q2, with a special emphasis on Q2, since it provides better results in simulations.

@@

We briefly introduce here the concept of convex regression, while we leave the main technical details to~\cite{Mazumder2019,simonetto_smooth_2021}. Suppose one has collected $\ell$ noisy measurements of a convex function $\varphi(\x): \R^n \to \R$ (say $y_i$) at points $\x_i \in \R^n$, $i \in I_{\ell}$, and (optionally) its gradients $\nabla_{\x} \varphi(\x_i)$. Then convex regression is a least-squares approach to estimate the generating function based on the measurements. Formally, letting $\varphi \in \mathcal{S}_{m,L}(\R^n)$, then one would like to solve the infinite-dimensional problem, 
\begin{equation}\label{eq.inf}
\hat{\varphi}_{\ell} \in \argmin_{\psi \in \mathcal{S}_{m, L}(\R^n)}\Big\{ \sum_{i\in I_{\ell}} (y_i - \psi(\x_i))^2 + \|\z_i - \nabla_{\x} \psi(\x_i)\|^2 \Big\} \,,
\end{equation}
where $y_i$ and $\z_i$ are the measurements of the function and the gradients at the data points.

The problem can be then equivalently decomposed into an estimation on the data points, to find the true function values and gradients at the data point ($\f = [\varphi_i]_{i \in I_{\ell}}$, $\bdelta = [\nabla \varphi_i]_{i \in I_{\ell}}$) which turns out to be a convex quadratically constrained quadratic program, 
\begin{subequations}\label{socp}
\begin{eqnarray}
(\f^*, \bdelta^*) &&\hskip-.5cm= \argmin_{\f\in\R^\ell\!,~ \bdelta \in\R^{n\ell}} \sum_{i\in I_{\ell}} (y_i - \varphi_i)^2 + \|\z_i - \bdelta_i \|^2 \\
\mathrm{s.t.:} &&
\varphi_i - \varphi_j - \bdelta_j^\transp (\x_i - \x_j) \geq \\ && \hskip-5mm \frac{1}{2(1 - m/L)}\left(\frac{1}{L}\|\bdelta_i - \bdelta_j \|^2_2 + m \|\x_i - \x_j\|_2^2 \right. \nonumber \\ && \hskip-5mm \left. - 2 \frac{m}{L} (\bdelta_j-\bdelta_i)^\transp (\x_j - \x_i) \right), \quad \forall i, j \in I_{\ell}. \nonumber
\end{eqnarray} 
\end{subequations}

@@

Then, to solve Q1, one could consider, for each time $t_k$, a number of evaluations of the function $f(\x; t_k)$ (and optionally its gradients) as noisy measurements of an underlying convex function $\hat{f}_k \in \mathcal{S}_{m,L}(\R^n)$ and using the least-squares approach above to \emph{project} function $f(\cdot; t_k)$ onto the space of ``good'' functions $\mathcal{S}_{m,L}(\R^n)$. Then, one can use the estimated function $\hat{f}_k$ and its estimated gradients to perform a step of the algorithm.

We include in the appendix all the technical details of such CvxReg-Boost algorithm, while we focus now on solving Q2 via operator regression, which is ultimately better performing.

\andrea{I think it's not super clear-- Add the algorithm here and the explanation of the compl. complexity and other in the appendix ? }

@@

A slightly different approach to convex regression is to look at the algorithm directly, instead of the function. First of all, we remind the reader that any algorithm can be seen as an operator that maps the current approximate optimizer $\x_k$ into a new approximate optimizer $\x_{k+1}$. For example, a gradient algorithm of the form,
\begin{equation}
\x_{k+1} = \x_k - \alpha \nabla_{\x} f(\x_k; t_k) \equiv \underbrace{(I - \alpha \nabla_{x}f_k)}_{=:\T_k} (\x_k)
\end{equation}
where $\T_k: \R^n \to \R^n$ is the gradient algorithm operator. Interpreting algorithms as operators (possibly averaged, monotone, etc.) has been extremely fruitful for characterizing their convergence properties~\cite{Rockafellar1976, Eckstein1989, bauschke_convex_2017,Ryu2015,Sherson2018}.

@@

First of all, using Fact~2.2 in \cite{ryu_operator_2020}, we know that an operator $\T$ is $\zeta$-contractive interpolable (and therefore extensible to the whole space) if and only if it satisfies
\begin{equation}
	\norm{\T \x_i - \T \x_j}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2, \quad \forall i,j \in I_{\ell}, \ i \neq j.
\end{equation}
Therefore, we can define the following convex quadratically constrained quadratic program as our regression problem:
\begin{equation}\label{eq:operator-regression}
\begin{split}
	\hat{\tv} &= \argmin_{ \R^{n\ell} \ni \tv = [\tv_i]_{i \in I_{\ell}}} \frac{1}{2} \sum_{i \in I_{\ell}} \norm{\tv_i - \T_k \x_i}^2 \\
	&\text{s.t.} \ \norm{\tv_i - \tv_j}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2 \ \forall i,j \in I_{\ell}, i \neq j,
\end{split}
\end{equation}
where again the cost function is a least square terms on the ``observations''' and the constraints enforce contractiveness. In particular, the optimal values $\hat{\tv}$ on the data points represent the evaluations of a $\zeta$-contracting operator when applied to those points, i.e., $\hat{\tv}_i = \hat{\T}_k\x_i$.

To extend the operator outside the data points in the whole space, we can use a 1945 results, and namely the Corollary to Theorem~5 of \cite{valentine_lipschitz_1945}. In particular, we are interested in interpolating a Lipschitz continuous operator (that is, non-expansive or contractive) from a set of evaluations. The data are the pairs $\{ (\x_i, \hat{\tv}_i) \}_{i \in I_{\ell}}$.

By the Corollary to Theorem~5 of \cite{valentine_lipschitz_1945}, we know that a $\zeta$-Lipschitz continuous map $\hat{\T}_k$ can be extended (that is, interpolated) in a new point $\x$, while preserving Lipschitz continuity $\zeta$, by the construction:
\begin{equation}\label{eq:interpolation}
	\hat{\T}_k \x = \begin{cases}
		\hat{\tv}_i & \text{if} \ \x = \x_i \\
		\hat{\tv} \in \bigcap_{i \in I_{\ell}} \ball{\hat{\tv}_i}{\zeta \norm{\x - \x_i}} & \text{otherwise}
	\end{cases}
\end{equation}
where $\ball{\cv}{r} \subset \R^n$ denotes a ball of center $\cv$ and radius $r$. That is, we can interpolate $\hat{\T}_k$ in a new point $\x$ by finding a point in the intersection of the balls centered in $\hat{\tv}_i$ with radius $\zeta \norm{\x - \x_i}$, $i = 1, \ldots, \ell$. Notice that the intersection $\bigcap_{i \in I_{\ell}} \ball{\hat{\tv}_i}{\zeta \norm{\x - \x_i}}$ is guaranteed to be non-empty by Theorem~1 of \cite{valentine_lipschitz_1945}.

By definition $\hat{\tv}$ satisfies
\begin{equation}
	\norm{\hat{\tv} - \hat{\tv_i}} \leq \zeta \norm{\x - \x_i}, \quad \forall i \in I_{\ell},
\end{equation}
which implies that Lipschitz continuity (and contractivity) is preserved.

\andrea{What happens to $\|\tv' - \tv''\|$ for $\x'$ and $\x''$ both interpolated? }

Interpolating a Lipschitz continuous operator requires therefore finding a point in the intersection of $\ell$ ball sets. In the Appendix [..], we show how to efficiently compute this intersection with the method of alternating projections~\cite{reich_projection_2015}.

@@

\andrea{Appendix?}

@@

The solution of the convex problem~\eqref{eq:operator-regression} can be obtained by off-the-shelf solvers, but generally the computational complexity can be a limiting factor, since the problem has a number of constraints that scales quadratically with the number of data points $\ell$. In particular, the computational complexity of interior-point methods would scale as $O()$. This is generally the case in non-parametric regression~\cite{.}.

Here, we propose a trivially parallel tailored-made algorithm that solves~\eqref{eq:operator-regression} more efficiently (formally .. [..]) based on Peaceman-Rachford splitting (PRS).

The idea is as follows: each pair of data points $i, j \in I_{\ell}$, $i \neq j$, gives rise to one constraint, for a total of $\ell(\ell-1)/2$ constraints. We define the following set of pairs
\begin{equation}
	\mathcal{V} = \left\{ e = (i,j) \ | \ i, j \in I_{\ell}, \ i < j \right\}
\end{equation}
which are ordered (that is, for example we take $(1,2)$ and not $(2,1)$, to avoid counting constraints twice). To each pair $e = (i,j)$ corresponds the constraint $\norm{\tv_i - \tv_j}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2$.

Let $\tv_{i,e}$ and $\tv_{j,e}$ be copies of $\tv_i$ and $\tv_j$ associated to the $e$-th constraint; then we can equivalently reformulate Problem~\eqref{eq:operator-regression} as
\begin{subequations}\label{eq:equivalent-problem}
\begin{align}
	&\min_{\tv_{i,e}, \tv_{j,e}} \frac{1}{2 (\ell-1)} \sum_{e \in \mathcal{V}} \norm{\begin{bmatrix} \tv_{i,e} \\ \tv_{j,e} \end{bmatrix} - \begin{bmatrix} \y_i \\ \y_j \end{bmatrix}}^2 \\
	&\text{s.t.} \ \norm{\tv_{i,e} - \tv_{j,e}}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2 \label{eq:interpolation-constraints} \\
	&\qquad \tv_{i,e} = \tv_{i,e'} \ \forall e, e' | i \sim e, e'. \label{eq:consensus-constraints}
\end{align}
\end{subequations}
Problem~\eqref{eq:equivalent-problem} is a strongly convex problem with convex constraints defined in the variables $\tv_{i,e}$.

Let $\xx$ be the vector stacking all the $\tv_{i,e}$, then the problem is equivalent to
\begin{align*}
	\min_{\xx} \psi(\xx) + \chi(\xx)
\end{align*}
where
$$
	\psi(\xx) = \frac{1}{2(\ell-1)} \norm{\xx - \y}^2 + \psi_1(\xx)
$$
with $\psi_1$ the indicator function imposing~\eqref{eq:interpolation-constraints} and $\xi$ the indicator function imposing the ``consensus'' constraints~\eqref{eq:consensus-constraints}. The problem can then be solved using the Peaceman-Rachford splitting (PRS) characterized by the following updates $\ell \in \mathbb{N}$:
\begin{subequations}
\begin{align}
	&\xx^j = \prox_{\rho \psi}(\z^j) \label{eq:prs-xi} \\
	&\vv^j = \prox_{\rho \chi}(2 \xx^j - \z^j) \\
	&\z^{j+1} = \z^j + \vv^j - \xx^j.
\end{align}
\end{subequations}
The proximal of $\chi$ corresponds to the projection onto the consensus space, and thus can be characterized simply by
$$
	\vv_{i,e}^j = \frac{1}{\ell-1} \sum_{e' | i \sim e'} \left( 2 \tv_{i,e'}^j - \z_{e'}^j\right).
$$
Regarding the proximal of $\psi$, $\psi$ is separable, in the sense that it can be written as
$$
	\psi(\xx) = \sum_{e \in \mathcal{V}} \left[ \frac{1}{2 (\ell-1)} \norm{\begin{bmatrix} \tv_{i,e} \\ \tv_{j,e} \end{bmatrix} - \begin{bmatrix} \y_i \\ \y_j \end{bmatrix}}^2 + \iota_{e}(\tv_{i,e},\tv_{j,e}) \right]
$$
where $\iota_{e}$ denotes the indicator function of~\eqref{eq:interpolation-constraints} \andrea{is it $\psi_1$?}. Therefore, the update~\eqref{eq:prs-xi} can be solved by solving (possibly in parallel) the problems
\begin{equation}\label{eq:local-updates}
\begin{split}
	(\tv_{i,e},\tv_{j,e}) &= \argmin_{\tv_{i,e},\tv_{j,e}} \left\{ \frac{1}{2 (\ell-1)} \norm{\begin{bmatrix} \tv_{i,e} \\ \tv_{j,e} \end{bmatrix} - \begin{bmatrix} \y_i \\ \y_j \end{bmatrix}}^2 + \right.\\ & \left.\qquad \frac{1}{2\rho} \norm{\begin{bmatrix} \tv_{i,e} \\ \tv_{j,e} \end{bmatrix} - \z_e^j}^2 \right\}\\
	&\text{s.t.} \quad \norm{\tv_{i,e} - \tv_{j,e}}^2 \leq \zeta^2 \norm{\x_i - \x_j}^2.
\end{split}
\end{equation}
Problems~\eqref{eq:local-updates} are convex QCQP with one constraint, and can be solved in the dual domain as scalar problems by using Newton's method~\cite{.}, yielding a total computational complexity of $O(..)$ for the solution of~\eqref{.}.

@@

Therefore, we are interested in learning a contracting $\hat{\T}_k$ by evaluations of $\T_k$. The OpReg-Boost algorithm can be described as follows.

\smallskip
\smallskip
\hrule
{\bf OpReg-Boost algorithm}
\smallskip
\hrule
{\bf Required:} number of points $\ell$, stepsize $\alpha$, contraction factor $\zeta$, initial condition $\x_0$.

At each time $t_k$ do:
\begin{enumerate}
	\item Sample a new problem, that is, observe $f(\x; t_k)$ and $g(\x; t_k)$;
	\item Learn the closest contracting operator to $\T_{k}$, say $\hat{\T}_k$ by:
	%%
	\begin{itemize}
		\item Choose $\ell$ points around $\x_k$ including $\x_k$ itself (e.g., adding a zero-mean Gaussian noise term)
		\item Evaluate the operator on the data points: $\tv_i = \T_{k} \x_i$, $i\in I_{\ell}$, i.e., $\tv_i = \x_i - \alpha \nabla_{\x} f(\x_i; t_k)$;
		\item Solve~\eqref{eq:operator-regression} with the PRS-based solver, and output $\hat{\tv}_k = \hat{\T}_k \x_k$,
\end{itemize}
	%%
	\item Apply $\x_{k+1} = \prox_{g_k} (\hat{\T}_k \x_k) = \prox_{g_k} (\hat{\tv}_k) $.
\end{enumerate}
\smallskip
\hrule

A couple of words are now in order. First, the computational complexity of the algorithm is dominated by the operation regression problem~\eqref{eq:operator-regression}, while its number of gradient calls is $\ell$-times the ones of a standard forward-backward algorithm.  At each time $t_k$, we perform $\ell$ gradient evaluations. We remark that $\ell$ can be as small as $5$ in practice.

Second, as one can see from step 3., there is no need for solving the interpolation problem~\eqref{eq:interpolation}, since the operation regression problem~\eqref{eq:operator-regression} already provides the evaluation of the regularized operator at the data point $\x_k$.

In an attempt to reduce gradient calls, we present next an interpolated version of OpReg-Boost.

@@

\andrea{Add intro}

\smallskip
\smallskip
\hrule
{\bf OpReg-Boost with interpolation algorithm}
\smallskip
\hrule
{\bf Required:} number of points $\ell$, stepsize $\alpha$, initial condition $\x_0$, interpolation steps $\tau$.

At each time $t_k$ do:
\begin{enumerate}
	\item Sample a new problem, that is, observe $f(\x; t_k)$ and $g(\x; t_k)$;
	\item {\bf IF} $k \,\textrm{mod}\, \tau = 0$ then, 
	\begin{enumerate}
	\item Learn the closest contracting operator to $\T_{k}$, say $\hat{\T}_k$ by:
	%%
	\begin{itemize}
		\item Choose $\ell$ points around $\x_k$ including $\x_k$ itself (e.g., adding a zero-mean Gaussian noise term)
		\item Evaluate the operator on the data points: $\tv_i = \T_{k} \x_i$, $i\in I_{\ell}$, i.e., $\tv_i = \x_i - \alpha \nabla_{\x} f(\x_i; t_k)$;
		\item Solve~\eqref{eq:operator-regression} with the PRS-based solver, and output $\hat{\tv}_k = \hat{\T}_k \x_k$,
\end{itemize}
\end{enumerate}
	{\bf ELSE},
	\begin{enumerate}
	\item Interpolate last available $\hat{\T}_t$ for $\tv_k = \x_k - \alpha \nabla_{\x} f(\x_k; t_k)$ by using~\eqref{eq:interpolation} with the method of alternating projections and output $\hat{\tv}_k$ 
\end{enumerate}
	%%
	\item Apply $\x_{k+1} = \prox_{g_k} (\hat{\tv}_k) $.
\end{enumerate}
\smallskip
\hrule

\andrea{add explanation and closure.}

@@

%------------------------------------------------------------------------------
\section{CvxReg-Boost}
